Election 2024

### 2024-11-03 paper/paper.qmd
Appendix B referencing

## Collection
Times/ Siena polls are conducted by telephone, using live interviews. About 96% of interviews were contacted by cellphone, and interviews were conducted in both English and spanish. In the Philadelphia polls occurring between September 11th and 16th, 240,000 calls were made to nearly 118,000 likely voters (https://www.nytimes.com/interactive/2024/09/19/us/politics/times-siena-inquirer-poll-pennsylvania-likely-electorate.html). This high call volume is key to maximising response from all groups. In a further effort to get a good sample additional effort is made to constant under-represented groups, for example people with lower education levels.

@@

## Methodology pros and cons
The Random contact method in which people are contacted does not provide random sampling, as non-response has made this method on it's own more redundant. (@bailey2023new). This is clearly a factor here as the response rate was below 2%. The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@Brick2013Nonresponse). The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generate more accurate. 
With such low response rates the possibility for risk, in spite of the effort made by the pollster, is that non-response and choice are correlated, this is known as non-response bias. It is not uncommon to assume that non-response should occur at random, especially as this poll documents repeatedly calling each likely voter, however in the case of elections there is previous evidence of a bias. Research into the previous two elections, and the polls proceeding then showed a high probability that non-responders were more likely to vote for Donald Trump, who was the Republican candidate in both 2016 and 2020 (@clinton2021task, @kennedy2018evaluation). With Trump running again in 2024, his popularity being underestimated is a distinct possibility, as is mentioned in the Discussion section.
The benefit of how the survey was taken is the detail added from the additional questions asked, some key questions allow for further stratification based on the answers. For example, survey information is gathered on self-reported likelihood to vote, and obviously this gives respondents answers far more weight if they are. One documented phenomenon worth mentioning is social desirability bias, that people do are more inclined to answer with more socially acceptable answers. This could occur throughout the questionnaire, however it is very likely that respondents overestimate their likelihood to vote (@social). The pollster’s use of both modeled turnout probabilities based on historical data and self-reporting improves the accuracy of predicting who is likely to vote dramatically. The poll’s use of likely voters still brings some uncertainty to its results, as the results may be impacted by non-voters, however siena/NYT took a methodology that aims to negate this as much as possible. Another benefit of live interviews is that clarification of answers is possible and the additional depth of the data that this provides. Qualitative insights to voter's choices or indecision creates a more robust and significant dataset. By comparison this is far more valuable than a singular binary choice quiz you may see online.

>>>>

## Collection
Times/ Siena polls are conducted by telephone, using live interviews. About 96% of interviews were contacted by cellphone, and interviews were conducted in both English and spanish. In the Philadelphia polls occurring between September 11th and 16th, 240,000 calls were made to nearly 118,000 likely voters (@nyt2024cross_tabs). This high call volume is key to maximising response from all groups. In a further effort to get a good sample additional effort is made to constant under-represented groups, for example people with lower education levels.

@@

## Methodology pros and cons
The Random contact method in which people are contacted does not provide random sampling, as non-response has made this method on it's own more redundant. (@bailey2023new). This is clearly a factor here as the response rate was below 2%. The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@Brick2013Nonresponse). The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generate more accurate. 
With such low response rates the possibility for risk, in spite of the effort made by the pollster, is that non-response and choice are correlated, this is known as non-response bias. It is not uncommon to assume that non-response should occur at random, especially as this poll documents repeatedly calling each likely voter, however in the case of elections there is previous evidence of a bias. Research into the previous two elections, and the polls proceeding then showed a high probability that non-responders were more likely to vote for Donald Trump, who was the Republican candidate in both 2016 and 2020 (@clinton2021task, @kennedy2018evaluation). With Trump running again in 2024, his popularity being underestimated is a distinct possibility, as is mentioned in the Discussion section.
The benefit of how the survey was taken is the detail added from the additional questions asked, some key questions allow for further stratification based on the answers. For example, survey information is gathered on self-reported likelihood to vote, and obviously this gives respondents answers far more weight if they are. One documented phenomenon worth mentioning is social desirability bias, that people do are more inclined to answer with more socially acceptable answers. This could occur throughout the questionnaire, however it is very likely that respondents overestimate their likelihood to vote (@socialdesire). The pollster’s use of both modeled turnout probabilities based on historical data and self-reporting improves the accuracy of predicting who is likely to vote dramatically. The poll’s use of likely voters still brings some uncertainty to its results, as the results may be impacted by non-voters, however siena/NYT took a methodology that aims to negate this as much as possible. Another benefit of live interviews is that clarification of answers is possible and the additional depth of the data that this provides. Qualitative insights to voter's choices or indecision creates a more robust and significant dataset. By comparison this is far more valuable than a singular binary choice quiz you may see online.
